\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

Throughout my doctoral studies, I have received a great deal of support and
assistance.

\bigbreak
\noindent I would first and foremost like to thank my supervisor, Prof. Dr.
Achim Tresch, whose expertise was invaluable in formulating my research
questions and methodology. His insightful feedback pushed me to sharpen my
thinking and brought my work to a higher level. He was instrumental in
defining the path of my research, and for his guidance through each stage of
my doctoral studies, I am immensely grateful.

\bigbreak
\noindent My heartfelt thanks also go out to the many excellent scientists
that collaborated with me on the publications included in my dissertation.

\bigbreak
\noindent I would also like to thank my colleagues at the Institute of Medical
Statistics and Computational Biology, and the Computational Biology workgroup
in particular, for their wonderful fellowship. The companionable atmosphere in
our office makes every day I spend with them a pleasure.

\bigbreak
\noindent Furthermore, without the support of Prof. Dr. Michael Melkonian,
Prof. Dr. Günter Plickert, and Dr. Lisa Stephan in my early course of
studies, I would not have made it this far.

\bigbreak
\noindent Moreover, I would like to thank my parents, who always believed in
my ability to be successful in academia. None of this would have been possible
without them and their unending encouragement, for which I am grateful beyond
words. I know that they could hardly be any prouder of me, but I hope they
take delight in knowing that my small contributions to the sum total of human
knowledge bear their subtle brushstrokes, too.

\bigbreak
\noindent Finally, I could not have completed my dissertation without the
support of all of my friends, especially Peter and Dave, who provided
invaluable support, stimulating discussions, as well as the essential
distraction to rest my mind outside of research.

% ┌────────────────────────────────────────────────────────────────────────────┐
% │ RRZK                                                                       │
% └────────────────────────────────────────────────────────────────────────────┘

\newpage\null\vfill
\sectionmark{Acknowledgements}
\noindent I thank the Regional Computing Center of the University of Cologne
(RRZK) for providing computing time on the DFG-funded (Funding number: INST
216\thinspace{}/\thinspace{}512\thinspace{}/\thinspace{}1FUGG) High
Performance Computing (HPC) system CHEOPS as well as support.
\vfill